Supporting Ecological Decision Making Using Feature-Selection and Variable Importance

### 2019-09-12 code/98-paper/ieee/ieee-manuscript.tex

\documentclass[letterpaper, journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

@@

\usepackage{booktabs}
\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

\subsection{Subsection Heading Here}
Subsection text here.

@@

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

@@

\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.5\textwidth] {defoliation-distribution-plot-1.pdf}
        \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
		\label{fig:defol-distr}
	\end{center}
\end{figure}

@@

\begin{figure} [t!]
	\begin{center}
\centering
		\includegraphics[width=0.5\textwidth] {study-area-hyperspectral.pdf}
		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
		\label{fig:study_area}
	\end{center}
\end{figure}

\begin{figure} [t!]
	\begin{center}
		\includegraphics[width=0.5\textwidth] {defol-grid-3000px.jpg}
        \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
		\label{fig:defol-trees}
	\end{center}
\end{figure}

@@

% filter methods
\begin{table}[b!]
\centering
\caption[t]{List of filter methods used in this work}
\begingroup\footnotesize
\begin{tabular}{lll}
	\\
	Name                                         & Group                             & Reference          \\
	\hline
	Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
	Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
	Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
	Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
	Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
	Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
\end{tabular}
\endgroup
\label{tab:filter-methods}
\end{table}

@@

% plot performance results
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.5\textwidth] {performance-results-1.pdf}
        \caption{Scored predictive performance (RMSE) of models across tasks. Prefix "CV" denotes that the learner was optimized using internal 10-fold CV while prefix "MBO" means that Bayesian optimization was used for hyperparameter optimization. The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on. Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = "Carscore", Info = "Information Gain", Borda = "Borda"). The second value in the attached label of each point shows the scored RMSE value of the respective setting.}
		\label{fig:perf-result}
	\end{center}
\end{figure}

% plot no filter vs best filter for each model and task
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.5\textwidth] {filter-effect-1.pdf}
        \caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}
		\label{fig:filter-effects}
	\end{center}
\end{figure}
\subsection{Permutation-based variable importance}

% plot no filter vs best filter for each model and task
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.5\textwidth] {filter-perf-all-1.pdf}
        \caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}
		\label{fig:filter-perf-all}
	\end{center}
\end{figure}

@@

\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.
Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
In our view using a buffer of 2 m is a good compromise between both worlds.

@@

\begin{figure} [h]
\begin{center}
\includegraphics[width=0.5\textwidth] {spectral-signatures-1.pdf}
\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
\label{fig:spectral-signatures}
\end{center}
\end{figure}

@@

\begin{figure} [h]
\begin{center}
\includegraphics[width=0.5\textwidth] {correlation-filter-nri-1.pdf}
\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
\label{fig:correlation-filters}
\end{center}
\end{figure}

\section{Effect of different \texttt{$n_{bins}$} values on filter "information gain"}

\begin{figure} [h]
\begin{center}
\includegraphics[width=0.5\textwidth] {correlation-nbins-1.pdf}
\caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}
\label{fig:correlation-nbins}
\end{center}
\end{figure}

>>>>

\documentclass[letterpaper, peerreview]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

@@

\usepackage{booktabs}
\usepackage{adjustbox}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
\IEEEpeerreviewmaketitle{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

% suppress "underfull hbox" warnings
\hbadness=99999

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
 \ifCLASSOPTIONpeerreview
 %\begin{center} \bfseries EDICS Category: 3-BBND \end{center}
 \fi

% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

@@

@@

\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
        \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
		\label{fig:defol-distr}
	\end{center}
\end{figure}

@@

\begin{figure} [t!]
	\begin{center}
\centering
		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
		\label{fig:study_area}
	\end{center}
\end{figure}

\begin{figure} [t!]
	\begin{center}
		\includegraphics[width=0.48\textwidth] {defol-grid-3000px.jpg}
        \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
		\label{fig:defol-trees}
	\end{center}
\end{figure}

@@

% filter methods
\begin{table}[b!]
\centering
\caption[t]{List of filter methods used in this work}
\begingroup\footnotesize
\begin{adjustbox}{width={0.48\textwidth},totalheight={\textheight},keepaspectratio}
\begin{tabular}{lll}
	\\
	Name                                         & Group                             & Reference          \\
	\hline
	Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
	Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
	Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
	Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
	Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
	Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
\end{tabular}
\end{adjustbox}
\endgroup
\label{tab:filter-methods}
\end{table}

@@

% plot performance results
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {performance-results-1.pdf}
        \caption{Scored predictive performance (RMSE) of models across tasks. Prefix "CV" denotes that the learner was optimized using internal 10-fold CV while prefix "MBO" means that Bayesian optimization was used for hyperparameter optimization. The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on. Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = "Carscore", Info = "Information Gain", Borda = "Borda"). The second value in the attached label of each point shows the scored RMSE value of the respective setting.}
		\label{fig:perf-result}
	\end{center}
\end{figure}

% plot no filter vs best filter for each model and task
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {filter-effect-1.pdf}
        \caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}
		\label{fig:filter-effects}
	\end{center}
\end{figure}
\subsection{Permutation-based variable importance}

% plot no filter vs best filter for each model and task
\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {filter-perf-all-1.pdf}
        \caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}
		\label{fig:filter-perf-all}
	\end{center}
\end{figure}

@@

\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.

Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
In our view using a buffer of 2 m is a good compromise between both worlds.

@@

\begin{figure} [ht]
\begin{center}
\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
\label{fig:spectral-signatures}
\end{center}
\end{figure}

@@

\begin{figure} [ht]
\begin{center}
\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
\label{fig:correlation-filters}
\end{center}
\end{figure}

\section{Effect of different \texorpdfstring{$n_{bins}$}{nbins} values on filter "information gain"}

\begin{figure} [ht]
\begin{center}
\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
\caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}
\label{fig:correlation-nbins}
\end{center}
\end{figure}
